Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-11 paper/paper.tex
results...

Table~\ref{tab:experimental-results} shows  that  LGC
improves the  garbage collection time  and efficiency for most  of the
programs.  Even for  programs where the time taken is  more it is very
competitive.  Memory usage  graphs are shown for  select benchmarks in
Figure~\ref{fig:memory-usage}.  As  the number of  garbage collections
tend  to  be  very large  our  graphs  show  only  a window  which  is
representative of the behavior for  that particular benchmark.  In all
the programs we can see that  the curve corresponding to LGC regularly
dips below the RGC curve.

One area  of concern is the  huge gap between the  actual liveness and
the liveness  percieved by  our collector.  In case  of LGC  for eager
languages  the gap  was  very  narrow and  almost  touched the  actual
liveness curve.  In case of  lazy languages, due to using reachability
for copying we  end up copying more  cells. Any data which  is part of
closure irrespective  of whether it is  live or not gets  copied. Thus
many non-live  cells get  copied during garbage  collection increasing
the drag  time.  Our  experiments have  shown that  unless we  use the
exact  execution  point  for  closures, bringing  down  the  number  of
non-live  cells  copied in  closures  cannot  be reduced.   Using  the
approximated liveness (liveness  at the creation point)  does not give
benefits  in terms  of non-live  cells inside  closures as  everything
would be live at the point of creation.

>>>>

Table~\ref{tab:experimental-results}  compares the  garbage collection
statisctics for  RGC and LGC  with reachability for  copying closures.
We report the  number of GC events, average number  of cells reclaimed
per  GC, average  number  of cells  touched per  GC,  total number  of
dragged cells  and total time  to perform  all collections.  It  is no
surprise  that  number  of  cells  reclaimed per  GC  is  higher  (and
consequently the  number of  GCs is lower)  for LGC  based collection.
LGC also reduces  the number of dragged  cells significantly. However,
the cost for LGC is that per  GC execution time is more for LGC, which
results in higher overall execution  time for garbage collection, even
with reduced  number of collections.   However, LGC run time  is still
competitive (slowdown within 5X of RGC  for worst case), and is better
for   3    benchmarks   (best    case   2X   speedup).     Note   that
\verb@gc_bench@~\cite{gc_bench}   is  a   synthestic  benchmark   that
allocates complete binary  trees of various sizes.  These binary trees
are  never used  by the  program. As  a result,  the benchmark  highly
favours LGC.   We have included  this benchmark for  completeness only
and the numbers for it  are not considered representative of practical
programs.

Memory   usage   graphs   for   select   benchmarks   in   are   shown
Figure~\ref{fig:memory-usage}, Column  (a).  As the number  of garbage
collections tend to  be very large, for each benchmark  we also show a
window which  is representative  of the  behavior for  that particular
benchmark (Column (b)).  In all the programs we can see that the curve
corresponding to  LGC (red  line) regularly dips  below the  RGC curve
(blue line), the dip is specially visible for the benchmarks where LGC
outperforms                      RGC                     substantially
(Table~\ref{tab:experimental-results}). The graphs  also include curve
for reachable cells  (black, obtained approximately by  forcing RGC to
run at  a very high frequency)  and the live cells  for particular run
(grey, obtained  by post processing  the heap access  at the end  of a
program).

One area  of concern is the  huge gap between the  actual liveness and
the liveness  percieved by our  collector.  In  case of LGC  for eager
languages~\ref{asati14lgc} the gap was  very narrow and almost touched
the actual  liveness curve.  In  case of  lazy languages, this  gap is
largely   due  to   approximate  reachability   based  collection   of
closures. To implement liveness based  collection of closures, we need
to  record liveness  data  for  each closure  and  update  it at  each
evaluation  point.   These  space  and  run  time  overheads  make  it
infeasible  to  incorporate  liveness  based  GC  for  closures  in  a
practical garbage collector.
